---
title: "Descriptive statistics"
author: ""
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(janitor)
library(ggthemes)
library(knitr)
library(kableExtra)


          
horses <- readxl::read_xlsx("../data/HR BEHAV database for Faith.xlsx", sheet="data") %>%
  filter(timepoint == 2 | timepoint==3 | timepoint == 4) %>%
  mutate(Group = case_when(stimulus == 1 ~ "cow",
                           stimulus == 2 ~ "lion"),
         approach_corrected = if_else(Approach_lat > Visible, Visible, Approach_lat),
         pct_latency = 100*approach_corrected/Visible)%>%
  as.data.frame()


library(sqldf)

#sqldf("SELECT DISTINCT stimulus, sex, age_class, count(1)/3 FROM horses group by stimulus, sex, age_class") %>%
#  View()

#sqldf("SELECT DISTINCT stimulus, age_class, count(1)/3 FROM horses group by stimulus, age_class") %>%
#  View()


```

## Boxplots of variables by timepoint and stimulus exposure Group 
 - Group A = exposed to cow stimulus, 
 - Group B = exposed to lion stimulus)
 


## SniffPC: Percent of time sniffing


```{r, echo=FALSE, warning=FALSE}

horses %>%
  group_by(timepoint, Group) %>%
  summarise(n=n(), mean=mean(SniffPC,na.rm=TRUE),
            SD = sd(SniffPC, na.rm=TRUE),
            median = median(SniffPC, na.rm = TRUE),
            .groups = 'keep') %>%
  as.data.frame() %>%
  kable(digits = 2) %>%
  kable_styling(full_width = F)


```

```{r, echo=FALSE, warning=FALSE}

ggplot(horses, aes(x=factor(timepoint), y=SniffPC,fill=Group)) +
  geom_boxplot(outlier.shape = NA, aes(fill=Group)) +
  geom_point(position = position_jitterdodge(), aes(fill=Group), pch=21, size =2) +
  theme_classic() +
  ylab ("Sniffing (%)") +
  xlab("Timepoint") +
  theme(text=element_text(family="Arial",  size=18),
        legend.justification = c(0.5,1),
        legend.title = element_blank())



```



## ChewPC - percent of time chewing


```{r, echo=FALSE, warning=FALSE}

horses %>%
  group_by(timepoint, Group) %>%
  summarise(n = n(), mean=mean(ChewPC,na.rm=TRUE),
            SD = sd(ChewPC, na.rm=TRUE),
            median = median(ChewPC, na.rm = TRUE),
            .groups = 'keep') %>%
  as.data.frame() %>%
  kable(digits = 2) %>%
  kable_styling(full_width = F)


```


```{r, echo=FALSE, warning=FALSE}


ggplot(horses, aes(x=factor(timepoint), y=ChewPC,fill=Group)) +
  geom_boxplot(outlier.shape = NA, aes(fill=Group)) +
  geom_point(position = position_jitterdodge(), aes(fill=Group), pch=21, size =2) +

  theme_classic() +
  ylab ("Chewing (%)") +
  xlab("Timepoint") +
  theme(text=element_text(family="Arial",  size=18),
        legend.justification = c(0.5,1),
        legend.title = element_blank())


```



## EarsPC: Percent of time with ears forward


```{r, echo=FALSE, warning=FALSE}

horses %>%
  group_by(timepoint, Group) %>%
  summarise(n = n(), mean=mean(EarsPC,na.rm=TRUE),
            SD = sd(EarsPC, na.rm=TRUE),
            median = median(EarsPC, na.rm = TRUE),
            .groups = "keep") %>%
  as.data.frame() %>%
  kable(digits = 2) %>%
  kable_styling(full_width = F)



```

```{r, echo=FALSE, warning=FALSE}

ggplot(horses, aes(x=factor(timepoint), y=EarsPC,fill=Group)) +
  geom_boxplot(outlier.shape = NA, aes(fill=Group)) +
  geom_point(position = position_jitterdodge(), aes(fill=Group), pch=21, size =2) +

  ylab("Ears forward (%)") +
  xlab("Timepoint") +
  theme_classic() +
  theme(text=element_text(family="Arial",  size=18),
        legend.justification = c(0.5,1),
        legend.title = element_blank())
```


## WithdrawPC: Percent of time withdrawing


```{r, echo=FALSE, warning=FALSE}

horses %>%
  group_by(timepoint, Group) %>%
  summarise(n = n(), mean=mean(WithdrPC,na.rm=TRUE),
            SD = sd(WithdrPC, na.rm=TRUE),
            median = median(WithdrPC, na.rm = TRUE),
            .groups="keep") %>%
  as.data.frame() %>%
  kable(digits = 2) %>%
  kable_styling(full_width = F)


```


```{r, echo=FALSE, warning=FALSE}

ggplot(horses, aes(x=factor(timepoint), y=WithdrPC,fill=Group)) +
  geom_boxplot(outlier.shape = NA, aes(fill=Group)) +
  geom_point(position = position_jitterdodge(), aes(fill=Group), pch=21, size =2) +
  theme_classic() +
  ylab("Withdrawal/retreat (%)") +
  xlab("Timepoint")+
  theme(text=element_text(family="Arial",  size=18),
        legend.justification = c(0.5,1),
        legend.title = element_blank())

```


## InteractPC: Percent of time interacting




```{r, echo=FALSE, warning=FALSE}

horses %>%
  group_by(timepoint, Group) %>%
  summarise(n = n(), mean=mean(InteractPC,na.rm=TRUE),
            SD = sd(InteractPC, na.rm=TRUE),
            median = median(InteractPC, na.rm = TRUE),
            .groups = "keep") %>%
  as.data.frame() %>%
  kable(digits = 2) %>%
  kable_styling(full_width = F)


```


```{r, echo=FALSE, warning=FALSE}

ggplot(horses, aes(x=factor(timepoint), y=InteractPC,fill=Group)) +
  geom_boxplot(outlier.shape = NA, aes(fill=Group)) +
  geom_point(position = position_jitterdodge(), aes(fill=Group), pch=21, size =2) +

  theme_classic() +
  ylab("Interacting with scent (%)") +
  xlab("Timepoint") +
  theme(text=element_text(family="Arial",  size=18),
        legend.justification = c(0.5,1),
        legend.title = element_blank())

```


## Blows_occur: total number of audible blows

```{r, echo=FALSE, warning=FALSE}

horses %>%
  group_by(timepoint, Group) %>%
  summarise(n = n(), mean=mean(Blows_occur,na.rm=TRUE),
            SD = sd(Blows_occur, na.rm=TRUE),
            median = median(Blows_occur, na.rm = TRUE),
            .groups="keep") %>%
  as.data.frame() %>%
  kable(digits = 2, caption = "Audible blows (instance)") %>%
  kable_styling(full_width = F)
 

```




```{r, echo=FALSE, warning=FALSE}

ggplot(horses, aes(x=factor(timepoint), y=Blows_occur,fill=Group)) +
  geom_boxplot(outlier.shape = NA, aes(fill=Group)) +
  geom_point(position = position_jitterdodge(), aes(fill=Group), pch=21, size =2) +

  theme_classic() +
  ylab("Audible blows (instances)") +
  xlab("Timepoint") +
  theme(text=element_text(family="Arial",  size=18),
        legend.justification = c(0.5,1),
        legend.title = element_blank())

```


## Approach latency (% time) 

```{r, echo=FALSE, warning=FALSE}

horses %>%
  group_by(timepoint, Group) %>%
  summarise(n = n(), mean=mean(pct_latency,na.rm=TRUE),
            SD = sd(pct_latency, na.rm=TRUE),
            median = median(pct_latency, na.rm = TRUE),
            .groups="keep") %>%
  as.data.frame() %>%
  kable(digits = 2, caption = "Approach latency (% time)") %>%
  kable_styling(full_width = F)
 

```




```{r, echo=FALSE, warning=FALSE}

ggplot(horses, aes(x=factor(timepoint), y=pct_latency,fill=Group)) +
  geom_boxplot(outlier.shape = NA, aes(fill=Group)) +
  geom_point(position = position_jitterdodge(), aes(fill=Group), pch=21, size =2) +
  theme_classic() +
  ylab("Approach latency (% time)") +
  xlab("Timepoint")+
  theme(text=element_text(family="Arial",  size=18),
        legend.justification = c(0.5,1),
        legend.title = element_blank())

```

## Mean Heart Rate

```{r, echo=FALSE, warning=FALSE}

horses %>%
  group_by(timepoint, Group) %>%
  summarise(n = n(), mean=mean(MeanHR,na.rm=TRUE),
            SD = sd(MeanHR, na.rm=TRUE),
            median = median(MeanHR, na.rm = TRUE),
            .groups="keep") %>%
  as.data.frame() %>%
  kable(digits = 2, caption = "Mean HR (BPM)") %>%
  kable_styling(full_width = F)
 

```




```{r, echo=FALSE, warning=FALSE}

ggplot(horses, aes(x=factor(timepoint), y=MeanHR)) +
  geom_boxplot(outlier.shape = NA, aes(fill=Group)) +
  geom_point(position = position_jitterdodge(), aes(fill=Group), pch=21, size =2) +
  theme_classic() +
  ylab("Mean HR (BPM)") +
  xlab("Timepoint")+
  theme(text=element_text(family="Arial",  size=18),
        legend.justification = c(0.5,1),
        legend.title = element_blank())

```
